Rotational Spectral Unmixing of Exoplanets:\\Degeneracies between Surface Colors and Geography

### 2017-04-20 paper/main.tex
adding figures

Unmixing the disk-integrated spectra of exoplanets provides a clue to \edit{heterogeneous surfaces} that we cannot directly resolve in the foreseeable future. 
It is particularly important for terrestrial planets with diverse surface compositions like Earth. 
Although previous work on unmixing the spectra of Earth from disk-integrated multi-band light curves appeared successful, here we \edit{point out} a mathematical degeneracy between the surface colors and their spatial  distributions. 
Nevertheless, \edit{useful constraints on the spectral shape of individual surface types may be obtained from the fact that albedo should be between 0 to 1, depending the orientation of the trajectory of the planet in color space}.  
We demonstrate the degeneracy and the possible constraints using both the mock data based on a toy model of Earth and the observed data of Earth. 
Despite the severe degeneracy, we are still able to recover an approximate albedo spectra for ocean. 
\edit{We discuss that observing at many wavelengths so that surfaces albedo is close to 0 or 1 at multiple wavelengths will provide better constraints, as well as the large covering fractions. }

@@

Direct imaging of exoplanets is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of Earth and other planets, as they are observed from an astronomical distance. 
\edit{Disk-integrated spectra or even multi-band photometry (colors) in principle include the effects of the surface reflectance as well as the atmosphere. }
%Atmospheric molecules are identifiable through spectral absorption features \citep[e.g.,][]{DesMarais2002} but surface reflectance spectra also affect the spectra, which could be measured through low-resolution spectra, or multi-band photometry. 
However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[b!]
    \begin{center}
\includegraphics[width=\hsize]{schematics_ver2.pdf}
    \end{center}
    \caption{\edit{Schematic figure to illustrate the relation between observed data ($\{{\bf d}_i\} $; crosses) and surface spectra ($\{{\bf s}_k\} $; filled points). }}
\label{fig:schematic}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Meanwhile, in (b) we swap the \edit{sand} and ocean.  
Although \edit{sand} is now the dominant component, 
%the true location of the soil is rather not preferred based on 
our analysis prefers other locations toward the top of the allowed region, because there is more space there. 
Therefore, large covering fraction is not a sufficient condition for obtaining strong constraints on the color of a surface. 
In other words, a surface spectrum can be well retrieved if it has favorable geography (ideally an entire hemisphere solely covered with that surface) {\it and} it has favorable albedo spectrum (albedo close to 0 or 1 at as many wavelengths as possible).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=\hsize]{raddata_2_norm_noreg.pdf}
%	\includegraphics[width=\hsize]{raddata_2_norm_Mj_PCs.pdf}
	\includegraphics[width=\hsize]{raddata_2_norm_spectra.pdf}
    \end{center}
    \caption{Upper panel: PC plane from EPOXI observation of June 2008 \citep{Livengood2011}. Lower panel: the spectra corresponding to the locations (A$\sim$J) shown in the upper panel. The ocean, \edit{sand}, and clouds appear to be close to A, D, and F, respectively. }
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this case, the trajectory of planetary color is far from the boundary because the color variations of the real Earth are muted by the cloud cover. 
Consequently, we do not see three peaks but a spectrum of the posterior probability distribution. 
Nevertheless, we are likely to find one surface type in the upper part of the allowed region, near label A \edit{, which may be interpreted as ocean}.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside2.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside1.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside0.pdf}
    \end{center}
     \end{minipage}
    \caption{Making the map coarser using HealPix. Left: 192 pixels, Middle: 48 pixels, Right: 12 pixels. }
\label{fig:lowresolution}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{s:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper, we revisited the problem of estimating the albedo spectra of major surface type and their distribution\edit{s} from disk-integrated \edit{colors} of exoplanets. 
We pointed out the inherit degeneracy which makes it impossible to find unique solutions. 
Despite the degeneracy, 
\edit{some useful constraints on the surface spectra may be obtained, depending on the orientation of the trajectories of the disk-integrated colors in the color space.  %
We demonstrated using both a simplified toy model of Earth and the observed data by EPOXI that such constraints let us narrow down the likely solutions for the surface types. %
We found that the keys to better estimates is to observe at the wavelengths where surface albedo are close to 0 or 1. %
We also discussed how the unknown geometries }

%\memoYF{Ending statement needed here.}

>>>>

Unmixing the disk-integrated spectra of exoplanets provides a clue to \edit{heterogeneous surfaces} that we cannot directly resolve in the foreseeable future. 
It is particularly important for terrestrial planets with diverse surface compositions like Earth. 
Although previous work on unmixing the spectra of Earth from disk-integrated multi-band light curves appeared successful, we \edit{point out} a mathematical degeneracy between the surface colors and their spatial  distributions. 
Nevertheless, \edit{useful constraints on the spectral shape of individual surface types may be obtained from the fact that albedo should be between 0 to 1, depending on the trajectory of the planet in color space}.  
We demonstrate the degeneracy and the possible constraints using both the mock data based on a toy model of Earth and the observed data of Earth. 
Despite the severe degeneracy, we are still able to recover an approximate albedo spectra for ocean. 
\edit{We find that observing at multiple wavelengths where surfaces albedo is close to 0 or 1 will provide better constraints. }

@@

Direct imaging of exoplanets is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of Earth and other planets, as they are observed from an astronomical distance. 
\edit{Disk-integrated spectra or even multi-band photometry (colors) in principle include the effects of the surface reflectance as well as the atmosphere. (We use the term spectra and colors interchangeably.) }
%Atmospheric molecules are identifiable through spectral absorption features \citep[e.g.,][]{DesMarais2002} but surface reflectance spectra also affect the spectra, which could be measured through low-resolution spectra, or multi-band photometry. 
However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[b!]
    \begin{center}
\includegraphics[width=0.8\hsize]{schematics_ver3.pdf}
    \end{center}
    \caption{\edit{Schematic figure to illustrate the relation between observed data ($\{{\bf d}_i\} $; crosses) and surface spectra ($\{{\bf s}_k\} $; filled points). The dashed cube shows the range where physical albedo can exist (between 0 and 1). }}
\label{fig:schematic}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Meanwhile, in (b) we swap the \edit{sand} and ocean.  
Although \edit{sand} is now the dominant component, 
%the true location of the soil is rather not preferred based on 
our analysis prefers other locations toward the top of the allowed region, because there is more space there. 
Therefore, large covering fraction is not a sufficient condition for obtaining strong constraints on the color of a surface. 
\edit{Better constraints surface spectra can be obtained if its location on the PC plane is close to more than one boundaries of the allowed region. }
In other words, a surface spectrum can be well retrieved if it has favorable geography (ideally an entire hemisphere solely covered with that surface) {\it and} it has favorable albedo spectrum (albedo close to 0 or 1 at as many wavelengths as possible).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=0.85\hsize]{PCs_raddata_2_norm.eps}
	\includegraphics[width=0.88\hsize]{raddata_2_norm_noreg_ver2.pdf}
%	\includegraphics[width=\hsize]{raddata_2_norm_Mj_PCs.pdf}
	\includegraphics[width=0.85\hsize]{raddata_2_norm_spectra.pdf}
    \end{center}
    \caption{Top panel: PCs from EPOXI June data. Middel panel: color contour similar to Figure \ref{fig:PCplane}, but based on EPOXI observation of June 2008 \citep{Livengood2011}. The red line shows the light curve trajectory. Bottom panel: the spectra corresponding to the locations (A$\sim$J) shown in the upper panel. The ocean, \edit{sand}, and clouds appear to be close to A, D, and F, respectively. }
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this case, the trajectory of planetary color is far from the boundary because the color variations of the real Earth are muted by the cloud cover. 
Consequently, we do not see three peaks but a spectrum of the posterior probability distribution. 
Nevertheless, we are likely to find one surface type in the upper part of the allowed region, near label A \edit{, which can be interpreted as ocean}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[hbt!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside2.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside1.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_PCplane_Nside0.pdf}
    \end{center}
     \end{minipage}
    \caption{Making the map coarser using HealPix. Left: 192 pixels, Middle: 48 pixels, Right: 12 pixels. }
\label{fig:lowresolution}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}
\label{s:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper, we revisited the problem of estimating the albedo spectra of major surface type and their distribution\edit{s} from disk-integrated \edit{colors} of exoplanets. 
We pointed out the inherit degeneracy which makes it impossible to find unique solutions. 
Despite the degeneracy, \edit{the following physical conditions narrow down the possible  solutions for the surface albedos:
the actual surface albedo spectra in the $J$-dimensional color space are (1) to be located on the PC (hyper-)plane, (2) to enclose all of the data points of the disk-integrated colors, and (3) to be in the $J$-dimensional hypercube, bounded by 0 and 1 for all axes. %
We demonstrated using both a simplified toy model of Earth and the observed data by EPOXI that such constraints point us to the approximate spectra of ocean. }

\edit{In more general, we find that the success of the estimates critically depends on the degree of excursions of the trajectory of the disk-integrated colors, and its orientation in the color space.  %
In other words, the estimates will be better if the covering fraction of each surface type becomes close to 1, as an intuition suggests. 
In addition, observations at the many wavelengths where surface albedo are close to 0 or 1 could improve the estimates. }
\edit{As the wavelength coverage/resolution increases, and as the light curves at more orbital phases are obtained, the performance of our procedure will be refined. [YF: too obvious??]}
